### 2019-03-28 submission/manuscript.tex
Incorporate figure legends

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\begin{document}

\vspace{35mm}

\hypertarget{fecal-short-chain-fatty-acids-are-not-predictive-of-colorectal-cancer-status-and-cannot-be-predicted-based-on-bacterial-community-structure}{%
\section{Fecal short chain fatty acids are not predictive of colorectal
cancer status and cannot be predicted based on bacterial community
structure}\label{fecal-short-chain-fatty-acids-are-not-predictive-of-colorectal-cancer-status-and-cannot-be-predicted-based-on-bacterial-community-structure}}

\vspace{35mm}

Marc A. Sze\({^1}\), Begüm Topçuoğlu\({^1}\), Nicholas A.
Lesniak\({^1}\), Mack T. Ruffin IV\({^2}\), Patrick D.
Schloss\({^1}\)\({^\dagger}\)

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

\(1\) Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI 48109

\(2\) Department of Family Medicine and Community Medicine, Penn State
Hershey Medical Center, Hershey, PA

\vspace{10mm}

\hypertarget{observation-format}{%
\subsection{Observation format}\label{observation-format}}

\newpage
\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

The gut microbiome is thought to have a role in the development of
colorectal cancer by protecting against and exacerbating inflammation.
Short chain fatty acids (SCFAs), including butyrate, have been shown to
have anti-inflammatory properties and are produced in large quantities
by colonic bacteria. We assessed whether there was an association
between fecal SCFA concentrations and the presence of colonic adenomas
or carcinomas in a cohort of individuals that was previously subjected
to 16S rRNA gene and metagenomic shotgun sequencing. We measured the
fecal concentrations of acetate, propionate, isobutyrate, and butyrate
within the cohort and found that there were no meaningful differences in
their concentration and disease status. When we incorporated these
concentrations into Random Forest classification models trained to
differentiate between people with normal colons and those with adenomas
or carcinomas, we found that they did not significantly improve the
ability of 16S rRNA gene or metagenomic gene sequence-based models to
classify individuals. Finally, we generated Random Forest regression
models trained to predict the concentration of each SCFA based on 16S
rRNA gene or metagenomic gene sequence data from the same samples. These
models performed poorly and were only able to explain less than 10\% of
the observed variation in the SCFA concentrations. These results support
the broader epidemiological data that questions the value of fiber
consumption for reducing the risks of colorectal cancer. Although it is
likely that bacterial metabolites may serve as biomarkers to detect
adenomas or carcinomas, fecal SCFA concentrations have limited value.

\newpage

\hypertarget{importance}{%
\subsection{Importance}\label{importance}}

Considering colorectal cancer is the third leading cancer-related cause
of death within the United States, there is a great need to detect
colorectal tumors early without invasive colonoscopy procedures and to
prevent the formation of tumors. Short chain fatty acids (SCFAs) are
often used as a surrogate for measuring gut health and for being
anti-carcinogenic because of their anti-inflammatory properties. We
evaluated the fecal SCFA concentration of a cohort of individuals with
varying colonic tumor burden who were previously analyzed to identify
microbiome-based biomarkers of tumors. We were unable to find an
association between SCFA concentration and tumor burden or use SCFAs to
improve our microbiome-based models of classifying people based on their
tumor status. Furthermore, we were unable to find an association between
the fecal community structure and SCFA concentrations. These data
indicate that there is not a meaningful link between the gut microbiome,
SCFAs, and tumor burden.

\newpage

Colorectal cancer is the third leading cancer-related cause of death
within the United States (1). Less than 10\% of cases can be attributed
to genetic risk factors (2). This leaves a significant role for
environmental, behavioral, and other factors such as smoking and diet
(3, 4). Colorectal cancer is thought to be initiated by a series of
mutations that accumulate as mutated cells proliferate leading to
adenomatous lesions, which are succeeded by carcinomas (2). Throughout
this progression, there are ample opportunities for bacterial
populations to create mutations, induce inflammation, and accelerate
tumorigenesis. Numerous studies in murine models have supported this
model (5--7). Additional cross sectional studies in humans have
identified microbiome-based biomarkers of disease (8). These studies
suggest that in some cases, it is the loss of bacterial populations that
produce short-chain fatty acids (SCFAs) that results in increased
inflammation and tumorigenesis.

SCFAs have have anti-inflammatory and anti-proliferative activities (9).
Manipulation of SCFAs in mouse models of colorectal cancer by direct
supplementation or feeding of fiber caused an overall reduction in tumor
burden (10). These results suggest that supplementation with substrates
that bacteria can ferment to produce SCFAs may confer beneficial effects
against colorectal cancer. Regardless, there is a lack of evidence that
increasing SCFA concentrations can protect against colorectal cancer in
humans. Based on similar observations, many microbiome studies use the
concentrations of SCFAs and the presence of 16S rRNA gene sequences from
organisms and the genes involved in producing them as a biomarker of a
healthy microbiota (11, 12). Case-control studies that have investigated
SCFA concentrations in colorectal cancer found that patients with
carcinomas had lower concentrations of SCFAs versus patients with
adenomas or individuals without colon tumors (13). Although this would
argue that increasing SCFA concentrations could be protective against
tumorigenesis, in randomized controlled trials fiber supplementation has
been inconsistently associated with protection against tumor formation
and recurrence (14, 15). These findings temper enthusiasm for treatments
that aim to use SCFAs as biomarkers or protection against tumorigenesis.

\textbf{SCFA concentrations do not meaningfully vary with diagnosis or
treatment.} To quantify the associations between colorectal cancer, the
microbiome, and SCFAs, we quantified the concentration of acetate,
propionate, isobutyrate, and butyrate in feces of previously
characterized individuals with normal colons (N=172) and those with
colonic adenomas (N=198) or carcinomas (N=120) (16). The only SCFA that
had a significantly different concentration across the diagnoses was
isobutyrate (P=0.0091; Figure 1A). Interestingly, the median
concentration of isobutyrate was 3.30 mmol/kg in people with normal
colons and it was 3.00 and 3.84 mmol/kg in people with colonic adenomas
or carcinomas, respectively. The difference in isobutyrate concentration
between people with adenomas and carcinomas was significantly different
(P=0.0065); however, the differences between people with normal colons
and those with adenomas or carcinomas was not significant (P=0.19 and
P=0.11). Among the subjects with adenomas and carcinomas, a subset
(N\textsubscript{adenoma}=41, N\textsubscript{carcinoma}=26) were
treated and sampled a year later (17). The only SCFA that changed
following treatment was isobutyrate, which decreased by 0.99 mmol/kg
(P=0.002; Figure 1B). For both the pre-treatment cross-sectional data
and the pre/post treatment data, we pooled the SCFA concentrations on a
per molecule of carbon basis and again failed to see any significant
differences (P\textgreater{}0.15). The low concentration of isobutyrate
relative to the other SCFAs, inconsistent concentrations, and unexpected
decrease in concentration with treatment makes it difficult to ascribe
much biological relevance to this observation.

\textbf{Combining SCFA and microbiome data does not improve the ability
to diagnose individual as having adenomas or carcinomas.} We previously
found that binning 16S rRNA gene sequence data into operational
taxonomic units based on 97\% similarity or into genera enabled us to
classify individuals as having adenomas or carcinomas using Random
Forest machine learning models (8, 16). We repeated that analysis but
added the concentration of the SCFAs as possible features to train the
models (Figure S1). Models trained using SCFAs to classify individuals
as having adenomas or carcinomas rather than normal colons had median
areas under the receiver operator characteristic curve (AUROC) that were
significantly greater than 0.5 (P\textsubscript{adenoma}\textless{}0.001
and P\textsubscript{carcinoma}\textless{}0.001); however, the AUROC
values to detect the presence of adenomas or carcinomas were only 0.54
and 0.55, respectively (Figure 2A). When we trained the models with the
SCFAs concentrations and operational taxonomic unit (OTU) or genus-level
relative abundances the AUROC values were not significantly different
from the models trained without the SCFA concentrations
(P\textgreater{}0.21; Figure 2A). These data demonstrate that knowledge
of the SCFA profile from a patient's fecal sample does not improve the
ability to diagnose a colonic lesion.

\textbf{Knowledge of microbial community structure does not predict SCFA
concentrations.} We next asked whether the fecal community structure was
predictive of fecal SCFA concentrations, regardless of a person's
diagnosis. We trained Random Forest regression models using 16S rRNA
gene sequence data binned into OTUs and genera to predict the
concentration of the SCFAs (Figure S2). The largest R\textsuperscript{2}
between the observed SCFA concentrations and the modeled concentrations
was 0.14 observed when using genus data to predict butyrate
concentrations (Figure 2B). We also used a smaller dataset of shotgun
metagenomic sequencing data generated from a subset of our cohort
(N\textsubscript{normal}=27, N\textsubscript{adenoma}=25, and
N\textsubscript{cancer}=26) (18). We binned genes extracted from the
assembled metagenomes into operational protein families (OPFs) or KEGG
categories and trained Random Forest regression models using metagenomic
to predict the concentration of the SCFAs (Figure S2). Similar to the
analysis using 16S rRNA gene sequence data, the metagenomic data was not
predictive of SCFA concentration; the largest amount of variation that
the models could explain was 0.055, which was observed when using KEGG
data to predict propionate concentrations (Figure 2B). Because of the
limited number of samples that we were able to generate metagenomic
sequence data from, we used our 16S rRNA gene sequence data to impute
metagenomes that were binned into metabolic pathways or KEGG categories
using PICRUSt (Figure S2). SCFA concentrations could not be predicted
based on the imputed metagenomic data. The largest amount of variation
that the models could explain was 0.097 observed when using KEGG data to
predict butyrate concentrations (Figure 2B). The inability to model SCFA
concentrations from microbiome data indicates that the knowledge of the
abundance of organisms and their genes was insufficient to predict SCFA
concentrations.

\textbf{Conclusion.} Our data indicate that fecal SCFA concentrations
are not associated with the presence of adenomas or carcinomas and that
they provide weak predictive power to improve the ability to diagnose
someone with one of these lesions. Furthermore, knowledge of the
taxonomic and genetic structure of gut microbiota was not predictive
SCFA concentrations. These results complement existing literature that
suggest that fiber consumption and the production of SCFAs are unable to
prevent the risk of developing colonic tumors. It is important to note
that our analysis concerned fecal SCFA concentrations and microbiome
characterization and that observations along the mucosa near the site of
lesions may provide a stronger association. Regardless, given the
growing literature in this area, it is unlikely that SCFAs are the
primary mechanism that limits tumorigenesis. This may be a cautionary
result to temper enthusiasm for SCFAs as a biomarker of gut health more
generally. Going forward it is critical to develop additional hypotheses
for how the microbiome and host interact to drive tumorigenesis to
better understand the disease process and identify biomarkers that will
allow early detection of tumors.

\newpage

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\textbf{Study design and sampling.} The overall study design and the
resulting sequence data have been previously described (16, 17). In
brief, fecal samples were obtained from 172 individuals with normal
colons, 198 individuals with colonic adenomas, and 120 individuals with
carcinomas. Of the individuals diagnosed as having adenomas or
carcinomas, a subset (N\textsubscript{adenoma}=41 and
N\textsubscript{carcinoma}=26) were sampled after treatment of the
lesion (median=255 days between sampling, IQR=233 to 334 days). Tumor
diagnosis was made by colonoscopic examination and histopathological
review of the biopsies (16). The University of Michigan Institutional
Review Board approved the studies that generated the samples and
informed consent was obtained from all participants in accordance to the
guidelines set out by the Helsinki Declaration.

\textbf{Measuring specific SCFAs.} The measurement of acetate,
propionate, isobutyrate, and butyrate used a previously published
protocol that used High-Performance Liquid Chromatography (HPLC) (19).
Two changes were made to the protocol. First, instead of using fecal
samples suspended in DNA Genotek OmniGut tubes, we suspended frozen
fecal samples in 1 mL of PBS. Second, instead of using the average
weight of fecal sample aliquots to normalize SCFA concentrations, we
used the actual weight of the fecal samples. These methodological
changes did not affect the range of concentrations of these SCFAs
between the two studies.

\textbf{16S rRNA gene sequence data analysis.} Sequence data from Baxter
et al.~(16) and Sze et al.~(17) were obtained from the Sequence Read
Archive (studies SRP062005 and SRP096978) and reprocessed using using
mothur v.1.42 (20). The original studies generated sequence data from V4
region of the 16S rRNA gene using paired 250 nt reads on an Illumina
MiSeq sequencer. The resulting sequence data were assembled into
contigs, screened to remove low quality contigs and chimeras. The
curated sequences were then clustered into OTUs at a 97\% similarity
threshold and assigned to the closest possible genus with an 80\%
confidence threshold trained on the reference collection from the
Ribosomal Database Project (v.16). We used PICRUSt (v.2.1.0-b) with the
recommended standard operating protocol to generate imputed metagenomes
based on the expected metabolic pathways, KEGG categories, and enzyme
commission numbers (21).

\textbf{Metagenomic DNA sequence analysis.} A subset of the samples from
the samples described by Baxter et al.~(16) were used to generate
metagenomic sequence data (N\textsubscript{normal}=27,
N\textsubscript{adenoma}=25, and N\textsubscript{cancer}=26). These data
were generated by Hannigan et al.~(18) and deposited into the Sequence
Read Archive (study SRP108915). Fecal DNA was subjected to shotgun
sequencing on an Illumina HiSeq using 125 bp paired end reads. The
archived sequences were already quality filtered and aligned to the
human genome to remove contaminating sequence data. We downloaded the
sequences and assembled them into contigs using MEGAHIT (22), which were
used to identify open reading frames (ORFs) using Prodigal (23). We
determined the abundance of each ORF by mapping the raw reads back to
the ORFs using Diamond (24). We clustered the ORFs into operational
protein families (OPFs) in which the clustered ORFs were more than 40\%
identical to each other using mmseq2 (25). We also used mmseq2 to map
the ORFs to the KEGG database and clustered the ORFs according to which
category the ORFs mapped.

\textbf{Random Forest models.} The classification models were built to
predict lesion type from microbiome information with or without SCFA
concentrations. The regression models were built to predict the SCFA
concentrations of acetate, butyrate, and propionate from microbiome
information. For classification and regression models, we pre-processed
the features by scaling them to vary between zero and one. Features with
no variance in the training set were removed from both the training and
testing sets. We randomly split the data into training and test sets so
that the training set consisted of 80\% of the full dataset while the
test set was composed of the remaining data. The training set was used
for hyperparameter selection and training the model, and the test set
was used for evaluating prediction performance. For each model, the best
performing hyperparameter, mtry, was selected in an internal five-fold
cross-validation of the training set with 100 randomizations. Six values
of mtry were tested and the value that provided the largest AUROC or
R\textsuperscript{2} was selected. We trained the random forest model
using the selected mtry value and predicted the held-out test set. The
data-split, hyperparameter selection, training and testing steps were
repeated 100 times to get a reliable and robust reading of model
prediction performance. We used AUROC and R\textsuperscript{2} as the
prediction performance metric for classification and regression models,
respectively. We used randomForest package implemented to the caret
package (version 4.6-14) in R statistical software (version 6.0-81) for
our models.

\textbf{Statistical analysis workflow.} Data summaries, statistical
analysis, and data visualizations were performed using R (v.3.5.1) with
the tidyverse package (v.1.2.1). To assess differences in SCFA
concentrations between individuals normal colons and those with adenomas
or carcinomas, we used the Kruskal-Wallis rank sum test. If a test had a
P-value below 0.05, we then applied a pairwise Wilcoxon rank sum test
with a Benjamini-Hochberg correction for multiple comparisons. To assess
differences in SCFA concentrations between individuals samples before
and after treatment we used paired Wilcoxon rank sum tests to test for
significance. To compare the median AUCROC for the held out data for the
model generated using only the SCFAs, we compared the distribution of
the data to the expected median of 0.5 using the Wilcoxon rank sum test
to test whether the model performed better than would be achieved by
randomly assigning the data to each diagnosis. When we compared the
Random Forest models generated without and with SCFA data included, we
used Wilcoxon rank sum tests to determine whether the models with the
SCFA data included did better.

\textbf{Code availability.} The code for all sequence curation and
analysis steps including an Rmarkdown version of this manuscript is
available at \url{https://github.com/SchlossLab/Sze_SCFACRC_XXXX_2019/}.

\newpage

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

The authors thank the Great Lakes-New England Early Detection Research
Network for providing the fecal samples that were used in this study. We
would thank the University of Michigan Center for Microbial Systems for
enabling our short-chain fatty acid analysis. Support for MAS came from
the Canadian Institute of Health Research and the National Institutes of
Health (UL1TR002240). Support for PDS came from the National Institutes
of Health (P30DK034933 and R01CA215574).

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Siegel2016}{}%
1. \textbf{Siegel RL}, \textbf{Miller KD}, \textbf{Jemal A}. 2016.
Cancer statistics, 2016. CA: A Cancer Journal for Clinicians
\textbf{66}:7--30.
doi:\href{https://doi.org/10.3322/caac.21332}{10.3322/caac.21332}.

\leavevmode\hypertarget{ref-Fearon1990}{}%
2. \textbf{Fearon ER}, \textbf{Vogelstein B}. 1990. A genetic model for
colorectal tumorigenesis. Cell \textbf{61}:759--767.
doi:\href{https://doi.org/10.1016/0092-8674(90)90186-i}{10.1016/0092-8674(90)90186-i}.

\leavevmode\hypertarget{ref-FlissIsakov2017}{}%
3. \textbf{Fliss-Isakov N}, \textbf{Zelber-Sagi S}, \textbf{Webb M},
\textbf{Halpern Z}, \textbf{Kariv R}. 2017. Smoking habits are strongly
associated with colorectal polyps in a population-based case-control
study. Journal of Clinical Gastroenterology 1.
doi:\href{https://doi.org/10.1097/mcg.0000000000000935}{10.1097/mcg.0000000000000935}.

\leavevmode\hypertarget{ref-Lee2015}{}%
4. \textbf{Lee J}, \textbf{Jeon JY}, \textbf{Meyerhardt JA}. 2015. Diet
and lifestyle in survivors of colorectal cancer. Hematology/Oncology
Clinics of North America \textbf{29}:1--27.
doi:\href{https://doi.org/10.1016/j.hoc.2014.09.005}{10.1016/j.hoc.2014.09.005}.

\leavevmode\hypertarget{ref-Zackular2013}{}%
5. \textbf{Zackular JP}, \textbf{Baxter NT}, \textbf{Iverson KD},
\textbf{Sadler WD}, \textbf{Petrosino JF}, \textbf{Chen GY},
\textbf{Schloss PD}. 2013. The gut microbiome modulates colon
tumorigenesis. mBio \textbf{4}:e00692--13--e00692--13.
doi:\href{https://doi.org/10.1128/mbio.00692-13}{10.1128/mbio.00692-13}.

\leavevmode\hypertarget{ref-DeStefanoShields2016}{}%
6. \textbf{Shields CED}, \textbf{Meerbeke SWV}, \textbf{Housseau F},
\textbf{Wang H}, \textbf{Huso DL}, \textbf{Casero RA}, \textbf{O'Hagan
HM}, \textbf{Sears CL}. 2016. Reduction of murine colon tumorigenesis
driven by Enterotoxigenic Bacteroides fragilis using cefoxitin
treatment. Journal of Infectious Diseases \textbf{214}:122--129.
doi:\href{https://doi.org/10.1093/infdis/jiw069}{10.1093/infdis/jiw069}.

\leavevmode\hypertarget{ref-Tomkovich2017}{}%
7. \textbf{Tomkovich S}, \textbf{Yang Y}, \textbf{Winglee K},
\textbf{Gauthier J}, \textbf{Mühlbauer M}, \textbf{Sun X},
\textbf{Mohamadzadeh M}, \textbf{Liu X}, \textbf{Martin P}, \textbf{Wang
GP}, \textbf{Oswald E}, \textbf{Fodor AA}, \textbf{Jobin C}. 2017.
Locoregional effects of microbiota in a preclinical model of colon
carcinogenesis. Cancer Research \textbf{77}:2620--2632.
doi:\href{https://doi.org/10.1158/0008-5472.can-16-3472}{10.1158/0008-5472.can-16-3472}.

\leavevmode\hypertarget{ref-Sze2018}{}%
8. \textbf{Sze MA}, \textbf{Schloss PD}. 2018. Leveraging existing 16S
rRNA gene surveys to identify reproducible biomarkers in individuals
with colorectal tumors.
doi:\href{https://doi.org/10.1101/285486}{10.1101/285486}.

\leavevmode\hypertarget{ref-OKeefe2016}{}%
9. \textbf{O'Keefe SJD}. 2016. Diet, microorganisms and their
metabolites and colon cancer. Nature Reviews Gastroenterology \&
Hepatology \textbf{13}:691--706.
doi:\href{https://doi.org/10.1038/nrgastro.2016.165}{10.1038/nrgastro.2016.165}.

\leavevmode\hypertarget{ref-Bishehsari2018}{}%
10. \textbf{Bishehsari F}, \textbf{Engen P}, \textbf{Preite N},
\textbf{Tuncil Y}, \textbf{Naqib A}, \textbf{Shaikh M}, \textbf{Rossi
M}, \textbf{Wilber S}, \textbf{Green S}, \textbf{Hamaker B},
\textbf{Khazaie K}, \textbf{Voigt R}, \textbf{Forsyth C},
\textbf{Keshavarzian A}. 2018. Dietary fiber treatment corrects the
composition of gut microbiota, promotes SCFA production, and suppresses
colon carcinogenesis. Genes \textbf{9}:102.
doi:\href{https://doi.org/10.3390/genes9020102}{10.3390/genes9020102}.

\leavevmode\hypertarget{ref-Sanna2019}{}%
11. \textbf{Sanna S}, \textbf{Zuydam NR van}, \textbf{Mahajan A},
\textbf{Kurilshikov A}, \textbf{Vila AV}, \textbf{Võsa U},
\textbf{Mujagic Z}, \textbf{Masclee AAM}, \textbf{Jonkers DMAE},
\textbf{Oosting M}, \textbf{Joosten LAB}, \textbf{Netea MG},
\textbf{Franke L}, \textbf{Zhernakova A}, \textbf{Fu J},
\textbf{Wijmenga C}, \textbf{McCarthy MI}. 2019. Causal relationships
among the gut microbiome, short-chain fatty acids and metabolic
diseases. Nature Genetics.
doi:\href{https://doi.org/10.1038/s41588-019-0350-x}{10.1038/s41588-019-0350-x}.

\leavevmode\hypertarget{ref-Meisel2016}{}%
12. \textbf{Meisel M}, \textbf{Mayassi T}, \textbf{Fehlner-Peach H},
\textbf{Koval JC}, \textbf{O'Brien SL}, \textbf{Hinterleitner R},
\textbf{Lesko K}, \textbf{Kim S}, \textbf{Bouziat R}, \textbf{Chen L},
\textbf{Weber CR}, \textbf{Mazmanian SK}, \textbf{Jabri B},
\textbf{Antonopoulos DA}. 2016. Interleukin-15 promotes intestinal
dysbiosis with butyrate deficiency associated with increased
susceptibility to colitis. The ISME Journal \textbf{11}:15--30.
doi:\href{https://doi.org/10.1038/ismej.2016.114}{10.1038/ismej.2016.114}.

\leavevmode\hypertarget{ref-Ohigashi2013}{}%
13. \textbf{Ohigashi S}, \textbf{Sudo K}, \textbf{Kobayashi D},
\textbf{Takahashi O}, \textbf{Takahashi T}, \textbf{Asahara T},
\textbf{Nomoto K}, \textbf{Onodera H}. 2013. Changes of the intestinal
microbiota, short chain fatty acids, and fecal pH in patients with
colorectal cancer. Digestive Diseases and Sciences
\textbf{58}:1717--1726.
doi:\href{https://doi.org/10.1007/s10620-012-2526-4}{10.1007/s10620-012-2526-4}.

\leavevmode\hypertarget{ref-Yao2017}{}%
14. \textbf{Yao Y}, \textbf{Suo T}, \textbf{Andersson R}, \textbf{Cao
Y}, \textbf{Wang C}, \textbf{Lu J}, \textbf{Chui E}. 2017. Dietary fibre
for the prevention of recurrent colorectal adenomas and carcinomas.
Cochrane Database of Systematic Reviews.
doi:\href{https://doi.org/10.1002/14651858.cd003430.pub2}{10.1002/14651858.cd003430.pub2}.

\leavevmode\hypertarget{ref-Gianfredi2018}{}%
15. \textbf{Gianfredi V}, \textbf{Salvatori T}, \textbf{Villarini M},
\textbf{Moretti M}, \textbf{Nucci D}, \textbf{Realdon S}. 2018. Is
dietary fibre truly protective against colon cancer? A systematic review
and meta-analysis. International Journal of Food Sciences and Nutrition
\textbf{69}:904--915.
doi:\href{https://doi.org/10.1080/09637486.2018.1446917}{10.1080/09637486.2018.1446917}.

\leavevmode\hypertarget{ref-Baxter2016}{}%
16. \textbf{Baxter NT}, \textbf{Ruffin MT}, \textbf{Rogers MAM},
\textbf{Schloss PD}. 2016. Microbiota-based model improves the
sensitivity of fecal immunochemical test for detecting colonic lesions.
Genome Medicine \textbf{8}.
doi:\href{https://doi.org/10.1186/s13073-016-0290-3}{10.1186/s13073-016-0290-3}.

\leavevmode\hypertarget{ref-Sze2017}{}%
17. \textbf{Sze MA}, \textbf{Baxter NT}, \textbf{Ruffin MT},
\textbf{Rogers MAM}, \textbf{Schloss PD}. 2017. Normalization of the
microbiota in patients after treatment for colonic lesions. Microbiome
\textbf{5}.
doi:\href{https://doi.org/10.1186/s40168-017-0366-3}{10.1186/s40168-017-0366-3}.

\leavevmode\hypertarget{ref-Hannigan2017}{}%
18. \textbf{Hannigan GD}, \textbf{Duhaime MB}, \textbf{Ruffin MT},
\textbf{Koumpouras CC}, \textbf{Schloss PD}. 2017. Diagnostic potential
\& the interactive dynamics of the colorectal cancer virome.
doi:\href{https://doi.org/10.1101/152868}{10.1101/152868}.

\leavevmode\hypertarget{ref-Venkataraman2016}{}%
19. \textbf{Venkataraman A}, \textbf{Sieber JR}, \textbf{Schmidt AW},
\textbf{Waldron C}, \textbf{Theis KR}, \textbf{Schmidt TM}. 2016.
Variable responses of human microbiomes to dietary supplementation with
resistant starch. Microbiome \textbf{4}.
doi:\href{https://doi.org/10.1186/s40168-016-0178-x}{10.1186/s40168-016-0178-x}.

\leavevmode\hypertarget{ref-Schloss2009}{}%
20. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

\leavevmode\hypertarget{ref-Langille2013}{}%
21. \textbf{Langille MGI}, \textbf{Zaneveld J}, \textbf{Caporaso JG},
\textbf{McDonald D}, \textbf{Knights D}, \textbf{Reyes JA},
\textbf{Clemente JC}, \textbf{Burkepile DE}, \textbf{Thurber RLV},
\textbf{Knight R}, \textbf{Beiko RG}, \textbf{Huttenhower C}. 2013.
Predictive functional profiling of microbial communities using 16S rRNA
marker gene sequences. Nature Biotechnology \textbf{31}:814--821.
doi:\href{https://doi.org/10.1038/nbt.2676}{10.1038/nbt.2676}.

\leavevmode\hypertarget{ref-Li2015}{}%
22. \textbf{Li D}, \textbf{Liu C-M}, \textbf{Luo R}, \textbf{Sadakane
K}, \textbf{Lam T-W}. 2015. MEGAHIT: An ultra-fast single-node solution
for large and complex metagenomics assembly via succinct de bruijn
graph. Bioinformatics \textbf{31}:1674--1676.
doi:\href{https://doi.org/10.1093/bioinformatics/btv033}{10.1093/bioinformatics/btv033}.

\leavevmode\hypertarget{ref-Hyatt2010}{}%
23. \textbf{Hyatt D}, \textbf{Chen G-L}, \textbf{LoCascio PF},
\textbf{Land ML}, \textbf{Larimer FW}, \textbf{Hauser LJ}. 2010.
Prodigal: Prokaryotic gene recognition and translation initiation site
identification. BMC Bioinformatics \textbf{11}:119.
doi:\href{https://doi.org/10.1186/1471-2105-11-119}{10.1186/1471-2105-11-119}.

\leavevmode\hypertarget{ref-Buchfink2014}{}%
24. \textbf{Buchfink B}, \textbf{Xie C}, \textbf{Huson DH}. 2014. Fast
and sensitive protein alignment using DIAMOND. Nature Methods
\textbf{12}:59--60.
doi:\href{https://doi.org/10.1038/nmeth.3176}{10.1038/nmeth.3176}.

\leavevmode\hypertarget{ref-Steinegger2017}{}%
25. \textbf{Steinegger M}, \textbf{Söding J}. 2017. MMseqs2 enables
sensitive protein sequence searching for the analysis of massive data
sets. Nature Biotechnology.
doi:\href{https://doi.org/10.1038/nbt.3988}{10.1038/nbt.3988}.

\newpage

\includegraphics{figure_1.png}

\textbf{Figure 1. SCFA concentrations did not vary meaningfully with
diagnosis of colonic lesions or with treatment for adenomas or
carcinomas.} (A) We measured the concentration of fecal SCFAs from
individuals with normal colons (N=172) or those with adenoma (N=198) or
carcinomas (N=120). (B) A subset of individuals diagnosed with adenomas
(N=41) or carcinomas (N=26) who underwent treatment were resampled a
year after the initial sampling; one extreme propionate value (124.4
mmol/kg) was included in the adenoma analysis but censored from the
visualization for clarity.

\newpage

\includegraphics{figure_2.png}

\textbf{Figure 2. SCFA concentrations do not improve models for
diagnosing the presence of adenomas, carcinomas, or all lesions and
cannot be reliably predicted from 16S rRNA gene or metagenomic sequence
data.} (A) The median AUROC for diagnosing individuals as having
adenomas or carcinomas using SCFAs was slightly better than than chance
(depicted by horizontal line at 0.50), but did improve performance of
the models generated using 16S rRNA gene sequence data. (B) Regression
models that were trained using 16S rRNA gene sequence, metagenomic, and
PICRUSt data to predict the concentrations of SCFAs performed poorly
(all median R\textsuperscript{2} values \textless{} 0.14). Regression
models generated using 16S rRNA gene sequence and PICRUSt data included
data from 490 samples and those generated using metagenomic data
included data from 78 samples.

\newpage

\includegraphics{figure_s1.png}

\textbf{Figure S1. Comparison of training and testing results for
classification models shows that the models are robust and are not
overfit.} Random Forest classification models were generated to
differentiate between subjects with normal colons and those with
adenomas or carcinomas using 16S rRNA gene sequence data that were
clustered into genera or OTUs with and without including the four SCFAs
as additional features. Random Forest classification models were
generated by partitioning the samples into a training set with 80\% of
the data and a testing set with the remaining samples. The training data
was used to fit the mtry hyperparameter using 100 five-fold cross
validations for six values of mtry. The mtry with the largest AUROC was
selected to train the model, which was then tested on the held out 20\%
of the samples.

\newpage

\includegraphics{figure_s2.png}

\textbf{Figure S2. Comparison of training and testing results for
regression models shows that the models are robust and are not overfit.}
Random Forest regression models were generated to predict the
concentration of each SCFA using each subjects' microbiome data
generated using 16S rRNA gene sequence and metagenomic sequence data.
These regression models were generated by partitioning the samples into
a training set with 80\% of the data and a testing set with the
remaining samples. The training data was used to fit the mtry
hyperparameter using 100 five-fold cross validations for six values of
mtry. The mtry with the largest R\textsuperscript{2} was selected to
train the model, which was then tested on the held out 20\% of the
samples.

\end{document}